A Streaming Algorithm for 2-Center with Outliers in High Dimensions

### 2015-05-03 paper.tex
Some more fixes + submission

\begin{abstract}
We study the Euclidean 2-center problem with outliers in the data stream model. 
Given a stream of points in arbitrary $d$ dimensions, the goal is to find two congruent balls 
of minimum radius covering all but $z$ points. 
We provide a $(1.8+\eps)$-approximation streaming algorithm for the problem, 
improving upon the previous $(4 + \eps)$-approximation algorithm available for the problem.
The space complexity and update time of our algorithm is $\poly(d, z, {1 \over \eps})$,
independent of the size of the stream.
\end{abstract}

@@

In this paper, we focus on the \emph{data stream} model of computation
in which only a single pass over input is allowed,
and we have only a limited amount of working space available.
%which is typically sublinear in the input size.
This model is specially useful for processing large data sets,
as it does not require the entire data set to be stored in memory.

@@

The Euclidean $k$-center problem has been extensively studied in the literature.
%There is a considerable amount of work on the geometric $k$-center problem.
If $k$ is part of the input,
the problem in known to be NP-hard in two and more dimensions~\cite{fowler1981optimal},
and is even hard to approximate to within a factor close to 2, 
unless P$\,=\,$NP~\cite{feder1988optimal}.
Factor-2 approximation algorithms are available for the problem 
in any dimension~\cite{gonzalez1985clustering,feder1988optimal}.
%The $k$-center problem without outliers (i.e., $z=0$) in general (Euclidean) space can not be approximated with a smaller factor than $2$ ($1.822$), unless $\mathbf{P}=\mathbf{NP}$~\cite{bern1996approximation}.
%Gonzalez~\cite{gonzalez1985clustering} showed a $2$-approximation algorithm for general case of the problem.
For small $k$ and $d$, %there are more efficient algorithms.
better solutions are available.
The 1-center problem in fixed dimensions is known to be LP-type 
and can be solved in $O(n)$ time~\cite{chazelle1996linear}.
For 2-center in the plane, the current best algorithm
runs in $\O(n \log^2 n \log^2 \log n)$ time~\cite{chan1999more}.

@@

\begin{figure}[t]
	\centering
	\includegraphics[width=13em]{figs/definition}
	\caption{An example of $2$-center with 6 outliers}
	\label{fig:definition}
	\vspace{-.5em}
\end{figure}

@@

\paragraph{Our result}
In this paper, we study the 2-center problem with outliers in high dimensions.
We present a streaming algorithm for the problem that achieves
an approximation factor of $1.8 + \eps$, for any $\eps > 1$,
%using $O(d^3z/\eps)$ space.
using $\poly(d, z, {1 \over \eps})$ space and update time.
This improves the current streaming algorithm available for the problem
which has an approximation factor of $4+\eps$.
%Our work not only improves the best previous algorithm available for the case of 2-center,
%but is also counterintuitive in the sense that its 
The approximation factor of our algorithm matches
that of the best streaming algorithm for the 2-center problem without outliers.
This is somewhat surprising, considering that 
the best current approximation factors for streaming $k$-center with and without outliers 
differ by a factor of $\sqrt{2}$ for $k=1$,
and by a factor of 2  for general~$k$. %the 1-center problem.
See \Cref{table:summary} for a comparison.

@@

To warm up, we present a simple $2$-approximation streaming algorithm
for the 1-center problem with outliers.
It utilized a parallelization technique that 
will be used extensively in the rest of the paper.
The pseudocode is provided in \Cref{alg:1-center}.
The algorithm receives as input a stream of points, $P$,
and the number of outliers, $z$.
It also assumes that the first point $p_1$ of the stream is non-outlier.
We will show later how to remove this assumption.
The algorithm returns a ball $B$ covering all but at most $z$ points of $P$.
%

@@

\begin{algorithm}
\caption{\sc 1-Center$(P, z)$} 
\algsetup{indent=1.5em}
\label{alg:1-center}
\begin{algorithmic}[1]
	%\STATE {$\triangleright$ $c$ is a non-outlier point in the optimal solution}
	\STATE $B \gets B(p_1, 0)$
	\STATE $Q \gets \emptyset$
	\FORALL {$p$ in $P$}
		\IF {$p \notin B$}
			\STATE insert $p$ into $Q$
			\IF {$|Q| = z + 1$}
				\STATE $q$ $\gets$ closest point to $c$ in $Q$
				\STATE remove $q$ from $Q$
				\STATE $B \gets B(c, \len{cq})$
			\ENDIF
		\ENDIF
	\ENDFOR
	\RETURN {B}
\end{algorithmic}
\end{algorithm}

@@

\begin{proof}
Let $B^*(c^*, r^*)$ be the optimal solution,
and $c = p_1$ be a non-outlier point in the optimal solution.
Since $c$ is covered by $B^*$,
for all points $p \in B^*$, we have
$\len{cp} \le \len{cc^*} + \len{c^*p} \le 2r^*$.
Among the $z+1$ points furthest from $c$,
there is at least one point $q$ which is not outlier, % in the optimal solution,
and therefore, is contained in $B^*$
(see \Cref{fig:1center}). 
Thus, $\len{cq} \le 2r^*$,
and hence, the ball $B(c, \len{cq})$ returned by Algorithm~\ref{alg:1-center} 
is a 2-approximation. 
\end{proof}

@@

\begin{enumerate}
\item [(a)] $\delta < \delta^*$ %$\delta$ is a lower bound of $\delta^*$
\item [(b)] if $c_2$ is set, then
\begin{enumerate}
	\item[1.] $r \leq \delta/2$
	\item[2.] $B(c_p, \radius{c_p}) \subseteq B_2(c_2, r)$
\end{enumerate}
\item [(c)] $B_1 \cap B_2^* = \emptyset$
\item [(d)]  if $\cp \in B_2^*$, then 
	\begin{enumerate}
		\item [1.] $2r^* \le r$
		\item [2.] $B_2 \cap B_1^* = \emptyset$
		\item [3.] all points in $\Buffer$ are outliers
	\end{enumerate}

@@

Invariant (b1): 
%When an appropriate candidate for $c_2$ is found by Algorithm~\ref{alg:case-1}, 
%$c_2$ is set to that candidate point,
%and $\rc$ is set to $\frac{2}{\alpha}\len{c_1c_2}$.
%After that, points are added to $B_2$ by function \textproc{AddToB$_2$}
%only if they are within $\rc$-radius of $c_2$. 
When $c_2$ is set by Algorithm~\ref{alg:case-1}, 
it is the $(z+1)$-furthest point from $c_1$ in a set $T \subseteq P$,
and $\rc$ is set to $\frac{2}{\alpha}\len{c_1c_2}$.
Let $q$ be the $(z+1)$-furthest point from $c_1$ in the stream at that moment.
Then $\len{c_1c_2} \leq \len{c_1q}$.
Assuming $\alpha \ge 16$, we have 
$\radius{c_2} \leq \frac{1}{6}\frac{\alpha \len{c_1 q}}{(\alpha + 4)} \leq \delta / 6$, 
and hence,  
$$
	r \leq (2 + \frac{2}{\alpha})\radius{c_2} \leq 3 \times \radius{c_2}  \leq \delta /2,
$$
which means that after increasing $r$ by \textproc{AddToB$_2$},
the invariant still holds.

Invariant (b2): If $|B_2| < \dz$, then $c_p = c_2$ and $r = \radius{c_2}$, 
and hence, $B_2 = B(c_p, \radius{c_p})$. 
When the size of $B_2$ reaches $\dz$, 
the central point $\cp$ moves to the centerpoint of $B_2$,
and $\rc$ is increased by a factor of $(2 + \frac{2}{\alpha})$. 
Because the centerpoint of $B_2$ locates in $B_2$, then $c_p \in B(c_2, \radius{c_2})$.
Thus, if $|B_2| \ge \dz$ then $\len{c_2 c_p} \leq \radius{c_2}$, and hence
$$
	\radius{c_p} \leq \frac{2 (\len{c_1 c_2} + \len{c_2 c_p})}{\alpha} \leq \radius{c_2}(1 + \frac{2}{\alpha}).
$$ 
Therefore, $B(c_p, \radius{c_p}) \subseteq B_2(c_2, \radius{c_2}(2 + \frac{2}{\alpha}))$.

Invariant (c): A point $p$ can be added to $B_1$ in two cases. % by \Cref{alg:case-1}. 
The first case is in function \textproc{AddToB$_1$}, 
where the point is added to $B_1$ only if it is within distance $\delta$ of the center $c_1$,  
%when $\len{pc_1} \leq \delta$, 
which by invariant (a), 
guaranties $\len{pc_1} < \delta^*$. Therefore, $p \not \in B_2^*$ in this case.

The second case is in \Cref{alg:case-1}, when the buffer overflows and $B_2$ is non-empty.
The algorithm takes one of the following actions depending on the size of $B_2$.
If $|B_2| < \dz $, then $c_2 = c_p$. \Cref{alg:case-1} add $c_2$ to $B_1$.
By contradiction assume that $c_p \in B_2^*$, then due to invariants (b1) and (d1), 
$2r^* \leq r \leq \delta/2 < \delta^*/2$. 
Therefore, by Lemma \ref{lem:2r}, there must be at most $z$ points outside $B_1 \cup B_2$,
which contradicts the overflow of the buffer. 
If $|B_2| \geq \dz$, then $c_p$ is the centerpoint of the first $\dz$ points currently in $B_2$. 
In this case, we add all points of $B_2$ to $B_1$. 
By invariant (b), $r \leq \delta/2 < \delta^*/2$. 
Therefore, By Lemma \ref{lem:center-point}, $c_p \in B_1^*$ or $c_p \in B_2^*$. 
By contradiction, assume that $c_p \in B_2^*$. 
In this case, due to invariants (d1) and (b2), 
$B_2$ covers $B(c_p, \radius{c_p})$ and $2r^* \leq r$. 
Therefore, Similar to the previous part, it contradicts the overflow of the buffer.

Invariant (d1): 
By Observation \ref{obs:c+4}, if $c_1 \in B_1^*$ and $c_p \in B_2^*$,
then $1 \leq \frac{\len{c_1 c_p}}{\delta^*} \leq \frac{\len{c_1 c_p}}{\alpha r^*}$,
and as a result, $2r^* \leq \radius{c_p}$. 
If $|B_2| < \dz$, then $c_p = c_2$, and By \Cref{alg:case-1}, $r = \radius{c_2}$, hence, $2r^* \leq r$.
If $|B_2| \ge \dz$, then similar to invariant (b), 
$$
	2r^* \leq \radius{c_p} \leq (1 + \frac{2}{\alpha})\radius{c_2} \leq (2 + \frac{2}{\alpha})\radius{c_2} = r.
$$

@@

%\paragraph{Streaming Model}
To implement our algorithm in the streaming model,
we maintain the sets $B_1$ and $B_2$ 
in a data structure that supports adding points,
and gives a $\beta$-approximation to 1-center with $z$ outliers. %, for $k=0,\dots,z$.
Moreover, we maintain a set $B_u = B_1 \cup B_2$ in a similar data structure.
Note that these data structures do not need to maintain all the points. 
They only need to have a buffer of size $\dz$ that keeps the most recently added points.

@@

\begin{proof}
Our algorithm for answering queries is provided in ~\Cref{sec:query}.
To maintain $B_1$, $B_2$, and $B_u$, we 
use the streaming algorithm of~\cite{zarrabi2009streaming,chan2014streaming},
%To maintain an approximation to 1-center with $z$ outliers,
%we use the streaming algorithm of~\cite{zarrabi2009streaming}
%which combined with the %current best streaming algorithm for 1-center~\cite{chan2014streaming},
%result of~\cite{chan2014streaming},
which has an approximation factor of $1.22 \times \sqrt{2} <1.8$.
%The algorithm of~\cite{zarrabi2009streaming}
The algorithm has $O(d^3z)$ space and $\poly(d,z)$ update time.
%which implies the same bounds for \Cref{alg:case-1}.
Since we need to run $z + 1$ instances of \Cref{alg:case-1} in parallel, 
the space and update time are multiplied by a factor of~$z$.
\end{proof}

@@

After all points in $P$ are processed, if we are in one of the four states in the current path,
then the obtained solution is added to the feasible solutions.
Otherwise, the solution is not feasible, and is abandoned as in the KA algorithm.
Finally, we return the best solution among all computed feasible solutions. 
Kim and Ahn~\cite{kim2014improved} proved that in all feasible solutions
computed this way,
the larger ball among $B_1$ and $B_2$ has radius at most $3/2r$,
provided $\delta^* \le \alpha r^*$.
Assuming that we have a good estimate $r$ satisfying
$1.2r^* \le r < (1.2 + 2\eps/3)r^*$,
we get the following result.

@@

\noindent
Putting all together, we get the following result
from Theorems~\ref{thm:far-cmplx} and~\ref{thm:close-cmplx}:

@@

\begin{algorithm}
\caption{\sc Query} 
\algsetup{indent=1.5em}
\label{alg:query}
\begin{algorithmic}[1]
	%\STATE solutions $\gets \set{\Call{MinCover}{B_1, B_2, \Buffer}}$
	\STATE solutions $\gets \set{}$

	\STATE candidates $\gets \Buffer \cup B_2$
	\IF {$\card{B_2} \ge \dz$}
		\STATE candidates $\gets \set{c_2} \cup \Buffer $
		\STATE $B_1 \gets B_1 \cup B_2$
	\ENDIF

@@

\begin{proof}
	By invariants (a) and (b1), we know that $r \leq \delta/2 < \delta^*/2$. 
	By Lemma~\ref{lem:center-point}, $c_p \in B_1^* \cap B_2^*$. 
	Since $c_p \not \in B_2^*$, then $c_p \in B_1^*$. 
	On the other hand, by Observation~\ref{obs:intersection}, 
	$B_2$ intersect at most one of $B_1^*$ and $B_2^*$. 
	Therefore, $B_2 \cap B_2^* = \emptyset$.
\end{proof}

\noindent
Our query algorithm works as follows. 
For each candidate point $c \in C$, \Cref{alg:query} 
constructs $B'_1(c_1, \max \{\delta, \radius{c} \})$ and $B'_2(c, \radius{c})$. 
If the candidate $c$ equals the current $c_2$, then, $B_1 = B'_1$. 
Since $\radius{c_2} \leq r \leq \delta/2$ by invariant (b1) and $B'_2 \subset B_2$,
we do not need to construct any new set. 
For $c \neq c_2$, we know that $B_1 \subset B'_1$, and hence,
we just need to see which points in $\Buffer \cap B_2$ are located in $B'_1$.
When $|B_2| \geq \dz$ and $c \neq c_2$, then it means that assumed $c_p \not \in B_2^*$. 
Therefore, by Lemma \ref{lem:B2dz}, $B_2$ can be added to $B'_1$ without contradicting invariant (c). 
So in this case, we just need to see which points of $\Buffer$ must be added to $B'_1$. 
\Cref{alg:query} use functions \textproc{AddToB$'_1$} and \textproc{AddToB$'_2$} for adding a point to $B'_1$ and $B'_2$ respectively. 
The definition of these functions are the same as \textproc{AddToB$_1$} and \textproc{AddToB$_2$},
except that, they add point to $B'_i$ instead of $B_i$ ($i = 1, 2$).

>>>>

\begin{abstract}
We study the 2-center problem with outliers in high-dimensional data streams. 
Given a stream of points in arbitrary $d$ dimensions, the goal is to find two congruent balls 
of minimum radius covering all but $z$ points. 
We provide a $(1.8+\eps)$-approximation streaming algorithm for the problem, 
improving upon the previous $(4 + \eps)$-approximation algorithm available for the problem.
The space complexity and update time of our algorithm is $\poly(d, z, {1 \over \eps})$,
independent of the size of the stream.
\end{abstract}

@@

In this paper, we focus on the \emph{data stream} model of computation
where only a single pass over input is allowed,
and we have only a limited amount of working space available.
%which is typically sublinear in the input size.
This model is in particular useful for processing large data sets,
as it does not require the entire data set to be stored in memory.

@@

The Euclidean $k$-center problem has been extensively studied in the literature.
%There is a considerable amount of work on the geometric $k$-center problem.
If $k$ is part of the input,
the problem in known to be NP-hard in two and more dimensions~\cite{fowler1981optimal},
and is even hard to approximate to within a factor better than $1.82$, 
unless P$\,=\,$NP~\cite{feder1988optimal}.
Factor-2 approximation algorithms are available for the problem 
in any dimension~\cite{gonzalez1985clustering,feder1988optimal}.
%The $k$-center problem without outliers (i.e., $z=0$) in general (Euclidean) space can not be approximated with a smaller factor than $2$ ($1.822$), unless $\mathbf{P}=\mathbf{NP}$~\cite{bern1996approximation}.
%Gonzalez~\cite{gonzalez1985clustering} showed a $2$-approximation algorithm for general case of the problem.
For small $k$ and $d$, %there are more efficient algorithms.
better solutions are available.
The 1-center problem in fixed dimensions is known to be LP-type 
and can be solved in $O(n)$ time~\cite{chazelle1996linear}.
For 2-center in the plane, the current best algorithm
runs in $\O(n \log^2 n \log^2 \log n)$ time~\cite{chan1999more}.

@@

\begin{figure}[t]
	\centering
	\includegraphics[width=13em]{figs/definition}
	\caption{An example of $2$-center with 6 outliers.}
	\label{fig:definition}
	\vspace{-.5em}
\end{figure}

@@

\paragraph{Our result}
In this paper, we study the 2-center problem with outliers in high dimensions.
We present a streaming algorithm for the problem that achieves
an approximation factor of $1.8 + \eps$, for any $\eps > 1$,
%using $O(d^3z/\eps)$ space.
using $\poly(d, z, {1 \over \eps})$ space and update time.
This improves the current streaming algorithm available for the problem
which has an approximation factor of $4+\eps$.
%Our work not only improves the best previous algorithm available for the case of 2-center,
%but is also counterintuitive in the sense that its 
The approximation factor of our algorithm matches
that of the best streaming algorithm for the 2-center problem without outliers.
This is somewhat surprising, considering that 
the current best approximation factors for streaming $k$-center with and without outliers 
differ by a multiplicative factor of $\sqrt{2}$ for $k=1$,
and by a factor of 2  for general~$k$. %the 1-center problem.
See \Cref{table:summary} for a comparison.

@@

To warm up, we present a simple $2$-approximation streaming algorithm
for the 1-center problem with outliers.
It utilized a parallelization technique that 
will be used extensively in the rest of the paper.
The pseudocode is provided in \Cref{alg:1-center}.
The algorithm receives as input a stream of points, $P$,
and the number of outliers, $z$.
It assumes that the first point $p_1$ of the stream is non-outlier.
We will show later how to remove this assumption.
The algorithm returns a ball $B$ covering all but at most $z$ points of $P$.
%

@@

\begin{algorithm}
\caption{\sc 1-Center$(P, z)$} 
\algsetup{indent=1.5em}
\label{alg:1-center}
\begin{algorithmic}[1]
	%\STATE {$\triangleright$ $c$ is a non-outlier point in the optimal solution}
	\STATE $c \gets$ the first point in $P$
	\STATE $B \gets B(c, 0)$
	\STATE $Q \gets \emptyset$
	\FORALL {$p$ in $P$}
		\IF {$p \notin B$}
			\STATE insert $p$ into $Q$
			\IF {$|Q| = z + 1$}
				\STATE $q$ $\gets$ closest point to $c$ in $Q$
				\STATE remove $q$ from $Q$
				\STATE $B \gets B(c, \len{cq})$
			\ENDIF
		\ENDIF
	\ENDFOR
	\RETURN {B}
\end{algorithmic}
\end{algorithm}

@@

\begin{proof}
Let $B^*(c^*, r^*)$ be the optimal solution,
and $c$ be a non-outlier point in the optimal solution.
Since $c$ is covered by $B^*$,
for all points $p \in B^*$, we have
$\len{cp} \le \len{cc^*} + \len{c^*p} \le 2r^*$.
Among the $z+1$ points furthest from $c$,
there is at least one point $q$ which is not outlier, % in the optimal solution,
and therefore, is contained in $B^*$
(see \Cref{fig:1center}). 
Thus, $\len{cq} \le 2r^*$,
and hence, the ball $B(c, \len{cq})$ returned by Algorithm~\ref{alg:1-center} 
is a 2-approximation. 
\end{proof}

@@

\begin{enumerate}
\item [(a)] $\delta < \delta^*$ %$\delta$ is a lower bound of $\delta^*$
\item [(b)] $r \leq \delta/2$
%\begin{enumerate}
%	\item[1.] $r \leq \delta/2$
%	\item[2.] $B(c_p, \radius{c_p}) \subseteq B_2(c_2, r)$
%\end{enumerate}
\item [(c)] $B_1 \cap B_2^* = \emptyset$
\item [(d)]  if $\cp \in B_2^*$, then 
	\begin{enumerate}
		\item [1.] $2r^* \le r$
		\item [2.] $B_2 \cap B_1^* = \emptyset$
		\item [3.] all points in $\Buffer$ are outlier
	\end{enumerate}

@@

Invariant (b): 
When $c_2$ is set by Algorithm~\ref{alg:case-1}, 
it is the $(z+1)$-furthest point from $c_1$ in a set $T \subseteq P$,
and $\rc$ is set to $\frac{2}{\alpha}\len{c_1c_2}$.
Let $q$ be the $(z+1)$-furthest point from $c_1$ in the stream at that moment.
Then $\len{c_1c_2} \leq \len{c_1q}$.
Assuming $\alpha \ge 16$, we have 
$\radius{c_2} \leq \frac{1}{6}\frac{\alpha \len{c_1 q}}{(\alpha + 4)} \leq \delta / 6$, 
and hence,  
$$
	r \leq (2 + \frac{2}{\alpha})\radius{c_2} \leq 3 \times \radius{c_2}  \leq \delta /2,
$$
which means that the invariant holds, 
even after increasing $r$ by function \textproc{AddToB$_2$}.

Invariant (c): 
We first claim that if $c_2$ is set, then
$B(c_p, \radius{c_p}) \subseteq B_2(c_2, r)$.
If $|B_2| < \dz$, then $c_p = c_2$ and $r = \radius{c_2}$, 
and hence, $B_2 = B(c_p, \radius{c_p})$. 
When the size of $B_2$ reaches $\dz$, 
the central point $\cp$ moves to the centerpoint of $B_2$,
and $\rc$ is increased by a factor of $(2 + \frac{2}{\alpha})$. 
Because the centerpoint of $B_2$ locates in $B_2$, then $c_p \in B(c_2, \radius{c_2})$.
Thus, if $|B_2| \ge \dz$ then $\len{c_2 c_p} \leq \radius{c_2}$, and hence
$$
	\radius{c_p} \leq \frac{2 (\len{c_1 c_2} + \len{c_2 c_p})}{\alpha} \leq \radius{c_2}(1 + \frac{2}{\alpha}).
$$ 
Therefore, $B(c_p, \radius{c_p}) \subseteq B_2(c_2, \radius{c_2}(2 + \frac{2}{\alpha}))$.

A point $p$ can be added to $B_1$ in two cases. % by \Cref{alg:case-1}. 
The first case is in function \textproc{AddToB$_1$}, 
where the point is added to $B_1$ only if it is within distance $\delta$ of the center $c_1$,  
%when $\len{pc_1} \leq \delta$, 
which by invariant (a), 
guaranties $\len{pc_1} < \delta^*$. Therefore, $p \not \in B_2^*$ in this case.

The second case is in \Cref{alg:case-1}, when the buffer overflows and $B_2$ is non-empty.
The algorithm takes one of the following actions depending on the size of $B_2$.
If $|B_2| < \dz $, then $c_2 = c_p$. \Cref{alg:case-1} add $c_2$ to $B_1$.
By contradiction assume that $c_p \in B_2^*$, 
then by invariants (b) and (d1), 
$2r^* \leq r \leq \delta/2 < \delta^*/2$. 
Therefore, by Lemma \ref{lem:2r}, there must be at most $z$ points outside $B_1 \cup B_2$,
which contradicts the overflow of the buffer. 
If $|B_2| \geq \dz$, then $c_p$ is the centerpoint of the first $\dz$ points currently in $B_2$. 
In this case, we add all points of $B_2$ to $B_1$. 
By invariant (b), $r \leq \delta/2 < \delta^*/2$. 
Therefore, By Lemma \ref{lem:center-point}, $c_p \in B_1^*$ or $c_p \in B_2^*$. 
Suppose by way of contradiction that $c_p \in B_2^*$. 
In this case, by invariant (d1) and the claim we proved first,
$B_2$ covers $B(c_p, \radius{c_p})$ and $2r^* \leq r$. 
Therefore, Similar to the previous part, it contradicts the overflow of the buffer.

Invariant (d1): 
By Observation \ref{obs:c+4}, if $c_1 \in B_1^*$ and $c_p \in B_2^*$,
then $1 \leq \frac{\len{c_1 c_p}}{\delta^*} \leq \frac{\len{c_1 c_p}}{\alpha r^*}$,
and as a result, $2r^* \leq \radius{c_p}$. 
If $|B_2| < \dz$, then $c_p = c_2$, and by \Cref{alg:case-1}, 
$r = \radius{c_2}$, and therefore, $2r^* \leq r$.
If $|B_2| \ge \dz$, then similar to invariant (b), 
$$
	2r^* \leq \radius{c_p} \leq (1 + \frac{2}{\alpha})\radius{c_2} 
	\leq (2 + \frac{2}{\alpha})\radius{c_2} = r.
$$

@@

%\paragraph{Streaming Model}

@@

\begin{proof}
Our algorithm for answering queries is provided in ~\Cref{sec:query}.
It uses the current partition $B_1$, $B_2$, and Buffer, to compute
an optimal solution to 2-center with $z$ outliers.
In the streaming model, we cannot afford keeping all the points of $B_1$ and $B_2$. 
%To implement our algorithm in the streaming model,
Therefore, we maintain the sets $B_1$ and $B_2$ 
in a data structure that supports adding points,
and gives a $\beta$-approximation to 1-center with $k$ outliers, for $k=0,\dots,z$.
Moreover, we maintain a set $B_u = B_1 \cup B_2$ in a similar data structure.
Note that these data structures do not need to maintain all the points. 
They only need to have a buffer of size $\dz$ to keep the most recently added points.

To maintain $B_1$, $B_2$, and $B_u$, we 
use the streaming algorithm of~\cite{zarrabi2009streaming,chan2014streaming},
%To maintain an approximation to 1-center with $z$ outliers,
%we use the streaming algorithm of~\cite{zarrabi2009streaming}
%which combined with the %current best streaming algorithm for 1-center~\cite{chan2014streaming},
%result of~\cite{chan2014streaming},
which provides an approximation factor of $1.22 \times \sqrt{2} <1.8$.
%The algorithm of~\cite{zarrabi2009streaming}
The algorithm uses $O(d^3z)$ space and has $\poly(d,z)$ update time.
%which implies the same bounds for \Cref{alg:case-1}.
Since we need to run $z + 1$ instances of \Cref{alg:case-1} in parallel, 
the space and update time are multiplied by a factor of~$z$.
\end{proof}

@@

After all points in $P$ are processed, if we are in one of the four states in the current path,
then the obtained solution is added to the feasible solutions.
Otherwise, the solution is not feasible, and is abandoned as in the KA algorithm.
Finally, we return the best solution among all computed feasible solutions. 
Kim and Ahn~\cite{kim2014improved} proved that in all feasible solutions
computed this way,
the larger ball among $B_1$ and $B_2$ has radius at most $3/2r$,
provided $\delta^* \le \alpha r^*$.
(Their proof is stated for $\alpha = 2$, 
but can be extended to any $\alpha \ge 2$.)
Assuming that we have a good estimate $r$ satisfying
$1.2r^* \le r < (1.2 + 2\eps/3)r^*$,
we get the following.

@@

\noindent
Theorems~\ref{thm:far-cmplx} and~\ref{thm:close-cmplx}
together yield the following main result of the paper.

@@

\begin{algorithm}
\caption{\sc Query} 
\algsetup{indent=1.2em}
\label{alg:query}
\begin{algorithmic}[1]
	%\STATE solutions $\gets \set{\Call{MinCover}{B_1, B_2, \Buffer}}$
	\STATE solutions $\gets \set{}$

	\STATE candidates $\gets B_2 \cup \Buffer$
	\IF {$\card{B_2} \ge \dz$}
		\STATE candidates $\gets \set{c_2} \cup \Buffer $
		\STATE $B_1 \gets B_1 \cup B_2$
	\ENDIF

@@

\begin{proof}
	By invariants (a) and (b), we know that $r \leq \delta/2 < \delta^*/2$. 
	By Lemma~\ref{lem:center-point}, $c_p \in B_1^* \cap B_2^*$. 
	Since $c_p \not \in B_2^*$, then $c_p \in B_1^*$. 
	On the other hand, by Observation~\ref{obs:intersection}, 
	$B_2$ intersect at most one of $B_1^*$ and $B_2^*$. 
	Therefore, $B_2 \cap B_2^* = \emptyset$.
\end{proof}

\noindent
Our query algorithm works as follows. 
For each candidate point $c \in C$, \Cref{alg:query} 
constructs $B'_1(c_1, \max \{\delta, \radius{c} \})$ and $B'_2(c, \radius{c})$. 
If the candidate $c$ equals the current $c_2$, then, $B_1 = B'_1$. 
Since $\radius{c_2} \leq r \leq \delta/2$ by invariant (b) and $B'_2 \subset B_2$,
we do not need to construct any new set. 
For $c \neq c_2$, we know that $B_1 \subset B'_1$, and hence,
we just need to see which points in $\Buffer \cap B_2$ are located in $B'_1$.
When $|B_2| \geq \dz$ and $c \neq c_2$, then it means that assumed $c_p \not \in B_2^*$. 
Therefore, by Lemma \ref{lem:B2dz}, $B_2$ can be added to $B'_1$ without contradicting invariant (c). 
So in this case, we just need to see which points of $\Buffer$ must be added to $B'_1$. 
\Cref{alg:query} use functions \textproc{AddToB$'_1$} and \textproc{AddToB$'_2$} for adding a point to $B'_1$ and $B'_2$ respectively. 
The definition of these functions are the same as \textproc{AddToB$_1$} and \textproc{AddToB$_2$},
except that, they add point to $B'_i$ instead of $B_i$ ($i = 1, 2$).
